OpenMPIR

### 2017-09-08 paper.tex
More tweaks.

Putting failures aside, performance for our implementation is quite strong,
generally outperforming or matching the best of existing implementations. For
benchmarks where the overhead to work ratio is high is where the Tapir
implementation really shines. For example, for the \texttt{Fibonacci}
benchmark, our implementation finishes significantly faster than any of the
others. While ICC failed for that sized input, we did test with smaller inputs
and found its similarly outpaced by the Tapir implementation.  \texttt{NQueens}
and \texttt{FFT} show similar behaviour, to lesser degrees. In contrast,
benchmarks that have a lower overhead to work ratio unsurprisingly differ less
in performance. For example,

@@

In this section we attempt to understand \emph{why} the performance varies in
the ways it does. While we can't hope to figure out every discrepancy in
performance, we can hope to get some ideas for why the performance of our
implementation is generally better, and where each of the implementations is
spending its time.

Our primary tool for this task will be performance counters. It's worth noting
that some information is difficult to infer from performance counters, such as
sources of contention in runtimes, reasons for memory locality issues, etc.
Still, it will give us some insight into performance bottlenecks.

Before we begin understanding why, it's worth noting again that the scope of the 
work that this paper is responsible for is only the frontend. All performance
gains due to the existing Tapir backend and Cilk runtime we sadly cannot claim
credit for.

We can break reasons for performance discrepancies into a few categories:

@@

While we would like to distinguish between these, it is difficult to do so with
only performance counter information. A more thorough investigation would
require deep knowledge of each of the frontends and runtimes. While the
importance of optimizations of parallel code is clear, we hypothesize that most
of the disparity present in these results is a function of runtime
implementation differences. As justification for this hypothesis, consider
the \texttt{Fibonacci} benchmark. There are no memory accesses to the heap, 
and the code leaves little room for optimization, algorithmic changes
notwithstanding. In other words, it is effectively a benchmark of how fast a
runtime can execute fork-join parallelism. \gsnote{still worth looking into the
optimization I mention in the implemtnation section.}

@@

This supports our hypothesis that most of the slowdown incurred is overhead in
the runtime. For \texttt{tapir}, the overhead is relatively small given the
size of the work chunks, for \texttt{clang} and \texttt{icc} it is
significantly larger, while for \texttt{gcc} it is completely overwhelming.

For the cases where performance is close between implementations, like
\texttt{sparselu}, we expect runtimes to be dominated by actual task work. This 
expectation is confirmed by investigating performance counters for \texttt{sparselu}:
every implementation spends roughly 90\% of it's time doing work in the task
bodies.

>>>>

Putting failures aside, performance for our implementation is quite strong,
generally outperforming or matching the best of existing implementations. For
benchmarks where the overhead-to-work ratio is high is where the Tapir (and 
Cilk) implementation shows significant benefit. For example, for the \texttt{Fibonacci}
benchmark, our implementation finishes significantly faster than any of the
others. While ICC failed for that sized input, we did test with smaller inputs
and found it was similarly outpaced by the Tapir implementation.  \texttt{NQueens}
and \texttt{FFT} show similar behaviour, to lesser degrees. In contrast,
benchmarks that have a lower overhead to work ratio unsurprisingly differ less
in performance. For example,

@@

In this section we speculate \emph{why} the performance varies in
the ways it does. While we have yet to figure out every discrepancy in
performance, we hope to get some ideas for why the performance of our
implementation is generally better and where each of the implementations are
spending a majority of the time. Our primary tool for this task will be performance 
counters. It's worth noting that some information is difficult to infer from 
performance counters, such as sources of contention, reasons for memory locality 
issues, etc. Still, they do give us some insight into performance bottlenecks.

Before we begin understanding why, it's worth noting again that the scope of the 
work that this paper is responsible for is only the frontend. All performance
gains due to the existing Tapir backend and Cilk runtime we sadly cannot claim
credit for. We can break reasons for performance discrepancies into a few categories:

@@

While we would like to distinguish between these, it is difficult to do so with
only performance counter information. A more thorough investigation would
require deep knowledge of each of the frontends and runtimes. While the
importance of optimizations of parallel code is clear, we hypothesize that most
of the disparity present in these results is more a function of runtime
implementation differences. As justification for this hypothesis, consider
the \texttt{Fibonacci} benchmark. There are no memory accesses to the heap, 
and the code leaves little room for optimization --- algorithmic changes
notwithstanding. In other words, it is effectively a benchmark of how fast a
runtime can execute fork-join parallelism. \gsnote{still worth looking into the
optimization I mention in the implemtnation section.}

@@

This supports our hypothesis that most of the slowdown incurred is overhead in
the runtime. For Tapir the overhead is relatively small given the
size of the work chunks, for \texttt{clang} and \texttt{icc} it is
significantly larger, and for \texttt{gcc} it is overwhelming.

For the cases where performance is close between the implementations, like
\texttt{sparselu}, we expect runtimes to be dominated by actual task work. This 
expectation is confirmed by investigating performance counters for \texttt{sparselu}:
every implementation spends roughly 90\% of it's time doing work in the task
bodies.
